---
title: "Data generation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data generation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

It is possible to calculate adherence without using generateChronicDrugExposure(). However guidelines are only provided for using with it.

## Working with cohorts

It is useful to restrict adherence calculations to only those individuals who belong to a specific cohort. AdherenceFromOMOP supports this directly through the `cohortSchema` and `cohortId` arguments in **generateChronicDrugExposure()**

```{r, eval = FALSE, include = TRUE}
# preparation, no need to replicate
cdm <- mockAdherenceFromOMOP()

cohort <- dplyr::tibble(
  cohort_definition_id = 1, subject_id = 1,
  cohort_start_date = as.Date("1980-01-01"),
  cohort_end_date = as.Date("2020-01-10")
)

DBI::dbWriteTable(conn = CDMConnector::cdmCon(cdm), name = DBI::Id(schema = "main", table = "cohort"), value = cohort, append = TRUE)
  
```

Only drug exposure events that satisfy the following conditions are included:

1.  The person_id is included in the cohort, and

2.  drug_exposure_start_date falls within the cohort start–end period

```{r, eval = FALSE}
chronicDrugExposureCohort <- generateChronicDrugExposure(
  cdm = cdm,
  name =  "test_table",
  overwrite = T,
  cohortSchema = "main",
  cohortId = 1,
)
```

## Using drugConceptIdList variable

You may optionally restrict the included drug exposure events to a specific list of **drug_concept_id** values using the `drugConceptIdList` argument. This list can be generated manually or, more conveniently, with **CodelistGenerator::getDrugIngredientCodes()**.

```{r, eval= FALSE}
ingredients <- c(...)
medicationGroups <- CodelistGenerator::getDrugIngredientCodes(cdm = cdm,
                                                      name = ingredients,
                                                      nameStyle = "{concept_code}")

ingredientNames <- names(medicationGroups)
drug_concept_id_lst <- unlist(medicationGroups, use.names = F)
```

