---
title: "Calculating Adherence"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Calculating Adherence}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Overview

Medication adherence calculations depend on the **CMA** (Continuous Medication Availability) measure selected. **AdherenceFromOMOP** supports CMA types **1–9**. For examples illustrating the behavior of different CMA types, see the [Intro to CMA](#0){style="font-size: 11pt;"} tutorial. In this package adherenceSlidingWindow and adherenceCalculations are wrappers for functions in [AdhereR](#0 "link to package"){style="font-size: 11pt;"} package.

For more detailed guidance, refer to the [AdhereR documentation](https://cran.r-project.org/web/packages/AdhereR/AdhereR.pdf "link to cran").

For large datasets, it is recommended to use `adherenceSlidingWindowLoader` and `adherenceCalculationsLoader`, which load data in chunks and write results directly to the database.

In [AdhereR](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0174426) software the follow-up window is the total period for which relevant medication events are recorded for patients, while the observation window is the period within the follow-up window over which adherence is actually computed.

Within the OMOP framework, adherence can be computed in two ways, depending on whether [cohorts](a02-generate-data.html) are used during data generation:

-   **When cohorts are present:**

    -   *Cohort start/end dates* define the **AdhereR observation window**.

    -   *Observation period start/end dates* define the **AdhereR follow-up window**.

-   **When cohorts are not used:**

    -   The *observation period start/end dates* define **both** the follow-up and observation windows.

## Medication groups

By default, CMA values are calculated using all medication data available in the input dataset. To compute adherence for specific drug groups or to separate medications into categories, the `medicationGroups` argument can be supplied. This argument expects a *named list*. It is recommended to generate these lists using the **CodelistGenerator** package.\
If medications overlap across groups, CMA values are unaffected (no data is omitted). For more details, see the `medication.groups` documentation in **AdhereR**.

Below is an example of manually creating medication groups:

```{r, eval = FALSE}
medicationGroup <- list(
  c(1361364),
  c(1830280)   # vectors of drug_concept_ids
)
groupNames <- c("group1", "group2")

names(medicationGroup) <- groupNames

# Optional: wrap for improved display
medicationGroup <- omopgenerics::newCodelist(medicationGroup)
```

## General medication adherence

This is used for calculating adherence throughout subjects observation period.

```{r, eval=FALSE, include=FALSE, message=FALSE}
cdm <- mockAdherenceFromOMOP()

chronicDrugExposure <- generateChronicDrugExposure(
  cdm = cdm,
  name = "chronic_drug_exposure_table",
  overwrite = T
)
```

```{r, eval=FALSE, include = TRUE, message = FALSE}
results <- adherenceCalculationsLoader(chronicDrugExposure, cdm, resultsTableName = "AdherenceFromOMOP_demo_cma5", cma = c("CMA5"), medicationGroup = medicationGroup)
results
```

## Medication adherence with sliding window

This approach computes adherence over a series of sliding windows within the observation period, allowing adherence to be examined over time. Unlike the AdhereR function output, window.ID values are not unique across all records for a given patient. They are assigned sequentially within each contiguous period without NA values and reset after gaps.

```{r, eval = FALSE, message = FALSE}
results <- adherenceSlidingWindowLoader(chronicDrugExposure, cdm, resultsTableName = "AdherenceFromOMOP_demo_sw", cma = c("CMA5"), delayObservationWindowStart = FALSE, medicationGroup = medicationGroup)
results
```

Setting `delayObservationWindowStart = TRUE` delays the start of observation period used for adherence calculations, in this case the sliding-window sequence until the first prescription is dispensed.

```{r, eval = FALSE}
results_delayed <- adherenceSlidingWindowLoader(
  chronicDrugExposure,
  cdm,
  resultsTableName = "AdherenceFromOMOP_demo_sw_delayed",
  cma = c("CMA5"),
  delayObservationWindowStart = TRUE,
  sliding.window.duration = 1,
  sliding.window.duration.unit = c("days", "weeks", "months", "years")[4],
  sliding.window.step.duration = 1,
  sliding.window.step.unit = c("days", "weeks", "months", "years")[4]
)
results_delayed
```
