---
title: "Example use cases with statins"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example use cases with statins}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

**Note:** Before running the code in this vignette, make sure you have installed and loaded all required packages. See the [Getting Started](a01-getting-started.html) vignette for installation instructions.

# Example use cases with statins

## Database connection

```{r, eval=FALSE}
readRenviron("~/.Renviron")

writePrefix <- "adherencefromomop_"

db <- DBI::dbConnect(
  RPostgres::Postgres(),
  dbname = Sys.getenv("DB_NAME"),
  host = Sys.getenv("DB_HOST"),
  user = Sys.getenv("DB_USERNAME"),
  password = Sys.getenv("DB_PASSWORD"),
  port = Sys.getenv("DB_PORT")
)

cdm <- CDMConnector::cdmFromCon(
  con = db,
  cdmSchema = Sys.getenv("OHDSI_CDM"),
  writeSchema = Sys.getenv("OHDSI_WRITE"),
  writePrefix = writePrefix
)
```

```{r, eval=FALSE}
statiinid <- c("Atorvastatin", "Rosuvastatin", "Simvastatin", "Pravastatin", "Fluvastatin", "Lovastatin", "Pitavastatin")
```

## Adherence for each statin group

```{r, eval=FALSE}
medicationGroups <- CodelistGenerator::getDrugIngredientCodes(
  cdm = cdm,
  name = statiinid,
  nameStyle = "{concept_name}"
)

ingredientNames <- names(medicationGroups)
concepts <- unlist(medicationGroups, use.names = F)
```

```{r, eval=FALSE}
# generate chronic drug exposure uses conceptSet for finding suitable drugs

all_statins <- generateChronicDrugExposure(
  cdm = cdm,
  conceptSet = medicationGroups,
  name = "statins_study",
  overwrite = T
)

# if the table already exists in database
all_statins <- dplyr::tbl(db, CDMConnector::inSchema(CDMConnector::cdmWriteSchema(cdm), table = "statins_study", CDMConnector::dbms(db)))
```

To calculate adherence for each medication group the groups must be defined first:

```{r, eval=FALSE}
adherenceCMA5 <- calculateAdherenceBatched(
  drugExposure = all_statins,
  cdm = cdm,
  medicationGroup = medicationGroups,
  name = "statins_cma5",
  cma = "CMA5"
)

# if the table already exists in database
adherenceCMA5 <- dplyr::tbl(db, CDMConnector::inSchema(CDMConnector::cdmWriteSchema(cdm), table = "statins_cma5", CDMConnector::dbms(db)))
```

## Adherence for each statin group using cohorts

For selecting people from specific cohorts, use a cohort table reference and cohort id values.

```{r, eval=FALSE}
# Get reference to cohort table
cohort_ref <- dplyr::tbl(db, CDMConnector::inSchema("ohdsi_results_202503", "cohort", CDMConnector::dbms(db)))

all_statins_cohort <- generateChronicDrugExposure(
  cdm = cdm,
  conceptSet = medicationGroups,
  name = "statins_study_1688",
  overwrite = T,
  cohortId = 1688,
  cohort = cohort_ref
)

# if the table already exists in database

all_statins_cohort <- dplyr::tbl(db, CDMConnector::inSchema(CDMConnector::cdmWriteSchema(cdm), table = "statins_study_1688", CDMConnector::dbms(db)))
```

### Combining medication groups for simvastatin and rosuvastatin

```{r, eval=FALSE}
medicationGroups <- CodelistGenerator::getDrugIngredientCodes(
  cdm = cdm,
  name = c("Atorvastatin", "Rosuvastatin", "Simvastatin", "Pravastatin", "Fluvastatin", "Lovastatin", "Pitavastatin"),
  nameStyle = "{concept_name}"
)

medicationGroups[["simvastatin_rosuvastatin"]] <- c(medicationGroups[["simvastatin"]], medicationGroups[["rosuvastatin"]])

medicationGroups[["simvastatin"]] <- NULL
medicationGroups[["rosuvastatin"]] <- NULL

names(medicationGroups)
```

```{r, eval=FALSE}
adherenceCMA5 <- calculateAdherenceBatched(
  drugExposure = all_statins_cohort,
  cdm = cdm,
  medicationGroup = medicationGroups,
  name = "statins_study_1688_medgroups",
  cma = "CMA5"
)
```

```{r, eval=FALSE}
adherenceCMA5SW <- calculateAdherenceSlidingWindowBatched(
  drugExposure = all_statins_cohort,
  cdm = cdm,
  medicationGroup = medicationGroups,
  name = "statins_study_1688_medgroups_sw",
  cma = "CMA5",
  # the following sets sliding window with duration of 3 years and step length of 3 years:
  sliding.window.duration = 3,
  sliding.window.duration.unit = c("days", "weeks", "months", "years")[4],
  sliding.window.step.duration = 3,
  sliding.window.step.unit = c("days", "weeks", "months", "years")[4]
)

adherenceCMA5SW <- dplyr::tbl(db, CDMConnector::inSchema(CDMConnector::cdmWriteSchema(cdm), table = "statins_study_1688_medgroups_sw", CDMConnector::dbms(db)))
```

## Add cohorts and data

Adding cohorts as features to CMA data requires cohort table as input variable. It is recommended to use CDMConnector for generating this.

```{r, eval=FALSE}
cohortSet <- CDMConnector::readCohortSet("./inst")
cdm <- CDMConnector::generateCohortSet(cdm, cohortSet, "cohort")

cohortTable <- dplyr::collect(cdm$cohort)

cma_values_with_cohort_features <- addCohorts(adherenceCMA5, cohortTable)
```

Adding body mass index is a separate function.

```{r, eval=FALSE}
cma_data_with_bmi <- addBMI(adherenceCMA5SW, cdm)
```

To get an overview this function reports the number of individuals in three groups: (1) individuals included in the cohort,

(2) individuals with calculated adherence values,

(3) individuals who have at least *n* drug exposure records within a given calendar year, calculated from the generated drug exposure table.

```{r, eval = FALSE}
stats <- summarisePatientCounts(
  cdm = cdm,
  adherenceData = adherenceCMA5SW,
  drugExposure = all_statins_cohort
)
```
